---
title: "Knowledge Weighted Estimate"
author: Asa Palley and Ville Satopää
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Knowledge Weighted Estimate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Introduction
The `metaggR` package implements the knowledge-weighted estimate proposed in [Palley and Satopää (2020)](https://papers.ssrn.com/sol3/Papers.cfm?abstract_id=3504286). 
This procedure aggregates judges' estimates of a continuous outcome. To do so, each judge is asked to provide two 
responses:

1. An estimate of the outcome, and
1. a prediction of the other judges' average estimate of the outcome.

Aggregation is done with the function `knowledge_weighted_estimate(E,P)` that inputs *J* judges’ estimates of the outcome (`E`) and predictions of others (`P`), and outputs the knowledge-weighted estimate. In the following two sections we will illustrate this aggregator on a simple example and on experimental data included in the package. 

# Example 1: Three Gorges Dam
This section illustrates `metaggR` on the Three Gorges Dam example in [Palley and Satopää (2020)](https://papers.ssrn.com/sol3/Papers.cfm?abstract_id=3504286). First, we will upload the library:

```{r setup}
library(metaggR)
```

Next, we can aggregate the judges' estimates:

```{r}
# Judges' estimates:
E = c(50, 134, 206, 290, 326, 374)
# Judges' predictions of others:
P = c(26, 92, 116, 218, 218, 206)
# Knowledge-weighted estimate:
knowledge_weighted_estimate(E,P)
```
Therefore the final knowledge-weighted estimate is 329.3266.

# Example 2: Calorie Counts
This section illustrates `metaggR` on real-world data collected in an experiment in Palley and Satopää (2020).
In this experiment participants were presented with 36 different pictures of food from different restaurants
and were asked to estimate the total number of calories in these dishes.
 Each response involves three steps:
 
1. __Initial Estimates__: On the first screen the participant was presented with a picture of a meal and asked _How many calories do you think are in this meal?_
1. __Predictions of Others__: On the second screen the participant saw the same picture, was reminded of their previous estimate, and given the statement: _We will be showing this picture to other participants as well. Just as we did with you, we will ask them how many calories they believe are in this meal_. The participant was then asked to predict _How many calories do you think that others  will guess on average?_
1. __Final Estimates__: On the third screen the participant saw the same picture again and was asked _After having reflected on others, what is y our own final best estimate of the number of calories in this meal?_

The `metaggR` package includes the data in the `calories` dataset. This dataset is a list with 4 elements:

1.  `true_calories`: A vector of true calorie counts of each 36 meals.
1. `estimates_initial`: A list of the judges' initial estimates of the calorie counts in each of the 36 meals.
1. `estimates_final`: A list of the judges' final estimates of the calorie counts in each of the 36 meals.
1. `predictions_of_others`: A list of the judges' predictions of the others' average estimate of the calorie counts in each of the 36 meals.

The elements of each member of `calories` correspond to the same meal. Specifically, the *j*th elements of `true_calories`, `estimates_initial`, `estimates_final`, and `predictions_of_others` represent the true calories, initial estimates, final estimates, and predictions of others of the *j*th meal.

To illustrate, we will consider the responses given for the first meal. First, we will load the package and the `calories` dataset:

```{r}
library(metaggR)
data(calories)
```

Next, we will pick out the responses given for the first meal:

```{r}
meal = 1
# True number of calories in the first meal:
(theta = calories$true_calories[meal])
# Judges' initial estimates of the number of calories in the first meal:
(E1 = calories$estimates_initial[[meal]])
# Judges' final estimates of the number of calories in the first meal:
(E2 = calories$estimates_final[[meal]])
# Judges' predictions of others' average estimate of the number of calories in the first meal:
(P = calories$predictions_of_others[[meal]])
```
A total of 73 judges provided responses for this meal. The true calorie count is 990 calories. The first judge under-estimated the calorie count and provided an initial and final estimates of 140 and 122 calories, respectively. This judge predicted that the others' average estimate is 112 calories. The root-mean-squared-errors (RMSE) of the initial and final estimates, and their knowledge-weighted estimates are:

```{r}
# RMSE of the initial estimates:
sqrt(mean((E1-theta)^2))
# RMSE of the final estimates:
sqrt(mean((E2-theta)^2))

# RMSE of the knowledge-weighted estimate based on judges' initial estimates:
(KWE1 =  knowledge_weighted_estimate(E1, P))
sqrt((KWE1 - theta)^2)
# RMSE of the knowledge-weighted estimate based on judges' final estimates:
(KWE2 =  knowledge_weighted_estimate(E2, P))
sqrt((KWE2 - theta)^2)

```
This shows that the knowledge-weighted estimate can improve the average accuracy of an individual judge. Specifically, based on the judges' initial and final estimates, the knowledge-weighted estimates are 814.64 and 865.07 calories, respectively. Given that the true calorie count is 990 calories, both aggregate estimates are too low but the knowledge-weighted estimate based on the judges' final estimates is more accurate. In this example, it improves the individual judges' RMSE from around 670 calories to around 125 calories.

